---
title: "Pv_Hep_scRNAseq_additional_file_3"
author: "Anthony Ruberto"
date: "`r Sys.Date()`"
output: html_document
---

Here, we will merge the P.vivax data. We will first merge data derived from replicate 1 and replicate 2 separately. Second, we will merge all data (replicate 1 and 2). The rationale for assessing the data from replicate 1 and 2 separately gives us an opportunity to pick up any biases that may arise that are biological or technical in nature. Ways these biases would manifest include differences in clustering outputs or differential transcript testing. 

```{r libraries}
library(BUSpaRse)
library(DropletUtils)
library(stringr)
library(tibble)
library(dplyr)
library(ggplot2)
library(scico)
library(scales)
library(rtracklayer)
library(Seurat)
library("leiden")
library(scater)
library(ggpubr)
library(viridis)
library(EnhancedVolcano)
library(pheatmap)
library(RColorBrewer)
```

Merge Day 5 and Day 9 post infection parasite data from replicate 1.
```{r merge replicate 1}
DefaultAssay(Hep51a.seu.viv.f1) <- "RNA"
Hep51a.seu.viv.f1<-NormalizeData(Hep51a.seu.viv.f1)
Hep51a.seu.viv.f1<-FindVariableFeatures(Hep51a.seu.viv.f1)
VariableFeaturePlot(Hep51a.seu.viv.f1)

DefaultAssay(Hep91a.seu.viv.f1) <- "RNA"
Hep91a.seu.viv.f1<-NormalizeData(Hep91a.seu.viv.f1)
Hep91a.seu.viv.f1<-FindVariableFeatures(Hep91a.seu.viv.f1)
VariableFeaturePlot(Hep91a.seu.viv.f1)

Hep59.1.seu<-merge(Hep51a.seu.viv.f1, Hep91a.seu.viv.f1)
DefaultAssay(Hep59.1.seu) <- "RNA"
Hep59.1.seu<-NormalizeData(Hep59.1.seu)
Hep59.1.seu<-FindVariableFeatures(Hep59.1.seu)
Hep59.1.seu<-ScaleData(Hep59.1.seu)

Hep59.1.seu  <- SCTransform(Hep59.1.seu , variable.features.n = NULL, 
                            variable.features.rv.th = 1.3)
VariableFeaturePlot(Hep59.1.seu)
Hep59.1.seu <- RunPCA(Hep59.1.seu , verbose = FALSE)
ElbowPlot(Hep59.1.seu)
Hep59.1.seu <- RunUMAP(Hep59.1.seu , dims = 1:20, verbose = FALSE)
Hep59.1.seu <- RunTSNE(Hep59.1.seu, dims = 1:20, perplexity = 20)
Hep59.1.seu <- FindNeighbors(Hep59.1.seu , dims = 1:20, verbose = FALSE)
Hep59.1.seu <- FindClusters(Hep59.1.seu, verbose = FALSE, resolution = 0.1, algorithm = 4)

Idents(Hep59.1.seu)<-"seurat_clusters"

ggarrange(
  DimPlot(Hep59.1.seu, label = F, split.by = "day", reduction = 'tsne', cols = c("#3C5488FF", "#DC0000FF"), group.by = "day") + NoLegend()+border(),
  DimPlot(Hep59.1.seu, label = F, reduction = 'tsne', cols = c("#00A087FF", "#4DBBD5FF")) + NoLegend()+border(),
  align = "hv", common.legend = F, ncol = 2, widths = c(2,1))

ggarrange(
  FeaturePlot(Hep59.1.seu, features = "nFeature_RNA", reduction = "tsne", dims = c(1,2))+border()+scale_color_viridis_c()+ggtitle(""),
  FeaturePlot(Hep59.1.seu, features = "nCount_SCT", reduction = "tsne", dims = c(1,2))+border()+scale_color_viridis_c()+ggtitle(""),
  FeaturePlot(Hep59.1.seu, features = "PVP01-0304700", reduction = "tsne")+border()+ggtitle("PVP01_0304700::liver specific protein 2")+ scale_color_viridis_c(),
  align = "hv", common.legend = F, ncol = 3)
```

Merge Day 5 and Day 9 post infection parasite data from replicate 2.
```{r merge replicate 2}
DefaultAssay(Hep54a.seu.viv.f1) <- "RNA"
Hep54a.seu.viv.f1<-NormalizeData(Hep54a.seu.viv.f1)
Hep54a.seu.viv.f1<-FindVariableFeatures(Hep54a.seu.viv.f1)
VariableFeaturePlot(Hep54a.seu.viv.f1)

DefaultAssay(Hep94a.seu.viv.f1) <- "RNA"
Hep94a.seu.viv.f1<-NormalizeData(Hep94a.seu.viv.f1)
Hep94a.seu.viv.f1<-FindVariableFeatures(Hep94a.seu.viv.f1)
VariableFeaturePlot(Hep94a.seu.viv.f1)

Hep59.2.seu<-merge(Hep54a.seu.viv.f1, Hep94a.seu.viv.f1)
DefaultAssay(Hep59.2.seu) <- "RNA"
Hep59.2.seu<-NormalizeData(Hep59.2.seu)
Hep59.2.seu<-FindVariableFeatures(Hep59.2.seu)
Hep59.2.seu<-ScaleData(Hep59.2.seu)

Hep59.2.seu  <- SCTransform(Hep59.2.seu , variable.features.n = NULL, 
                            variable.features.rv.th = 1.3)
VariableFeaturePlot(Hep59.2.seu)
Hep59.2.seu <- RunPCA(Hep59.2.seu , verbose = FALSE)
ElbowPlot(Hep59.2.seu)
Hep59.2.seu <- RunUMAP(Hep59.2.seu , dims = 1:20, verbose = FALSE)
Hep59.2.seu <- RunTSNE(Hep59.2.seu, dims = 1:20, perplexity = 15)
Hep59.2.seu <- FindNeighbors(Hep59.2.seu , dims = 1:20, verbose = FALSE)
Hep59.2.seu <- FindClusters(Hep59.2.seu, verbose = FALSE, resolution = 0.1, algorithm = 4)

Idents(Hep59.2.seu)<-"seurat_clusters"

ggarrange(
  DimPlot(Hep59.2.seu, label = F, split.by = "day", reduction = 'tsne', cols = c("#3C5488FF", "#DC0000FF"), group.by = "day") + NoLegend()+border(),
  DimPlot(Hep59.2.seu, label = F, reduction = 'tsne', cols = c("#00A087FF", "#4DBBD5FF")) + NoLegend()+border(),
  align = "hv", common.legend = F, ncol = 2, widths = c(2,1))

ggarrange(
  FeaturePlot(Hep59.2.seu, features = "nFeature_RNA", reduction = "tsne", dims = c(1,2))+border()+scale_color_viridis_c()+ggtitle(""),
  FeaturePlot(Hep59.2.seu, features = "nCount_SCT", reduction = "tsne", dims = c(1,2))+border()+scale_color_viridis_c()+ggtitle(""),
  FeaturePlot(Hep59.2.seu, features = "PVP01-0304700", reduction = "tsne")+border()+ggtitle("PVP01_0304700::liver specific protein 2")+ scale_color_viridis_c(),
  align = "hv", common.legend = F, ncol = 3)
```

Save RDS files.
```{r save RDS files}
saveRDS(Hep59.1.seu, "outputs/Hep59.1.seu.rds")
saveRDS(Hep59.2.seu, "outputs/Hep59.2.seu.rds")
```

Now that we have the replicates processed separately, let's combine the 2 to generate a Seurat object that we will use for downstream analyses. The processed data can be loaded in to start the analyses from here. 
```{r read in RDS rep 1 and rep 2}
Hep59.1.seu <- readRDS("outputs/Hep59.1.seu.rds")
Hep59.2.seu <- readRDS("outputs/Hep59.2.seu.rds")
```

Merge the replicates.
```{r merge rep 1 and rep 2}
Hep59.1.2.seu<-merge(Hep59.1.seu, Hep59.2.seu)
DefaultAssay(Hep59.1.2.seu) <- "RNA"
Hep59.1.2.seu  <- SCTransform(Hep59.1.2.seu , variable.features.n = NULL, 
                              variable.features.rv.th = 1.3)

Hep59.1.2.seu <- RunPCA(Hep59.1.2.seu , verbose = FALSE)
ElbowPlot(Hep59.1.2.seu)
Hep59.1.2.seu <- RunUMAP(Hep59.1.2.seu , dims = 1:20, verbose = FALSE)
Hep59.1.2.seu <- RunTSNE(Hep59.1.2.seu , dims = 1:20, verbose = FALSE, 
                         perplexity = 20)
```

Let's visualize the data.
```{r visualize rep 1 and rep 2 merged}
Idents(Hep59.1.2.seu)<-"seurat_clusters"

# TSNE - Day 5 , 9, clustering (Schizonts, Hypnozoites)
ggarrange(
  
  DimPlot(Hep59.1.2.seu, label = F, split.by = "day", reduction = 'tsne', 
          cols = c("#3C5488FF", "#DC0000FF"), group.by = "day") + NoLegend()+ border(),
  
  DimPlot(Hep59.1.2.seu, label = F, reduction = 'tsne', 
          cols = c("#00A087FF", "#4DBBD5FF")) + NoLegend()+border(),
  
  align = "hv", common.legend = F, ncol = 2, widths = c(2,1)
)

# TSNE - Day 5 , 9 overlaid
DimPlot(Hep59.1.2.seu, label = F, reduction = 'tsne', 
        cols = c("#3C5488FF", "#DC0000FF"), group.by = "day") + NoLegend()+border()

# TSNE - Day 5 , 9 overlaid, split by replicate

DimPlot(Hep59.1.2.seu, label = F, split.by = "replicate",reduction = 'tsne', 
        cols = c("#3C5488FF", "#DC0000FF"), group.by = "day") + NoLegend()+border()

# TSNE - nFeatureRNA, nCountRNA, LISP2
ggarrange(
  
FeaturePlot(Hep59.1.2.seu, features = "nFeature_RNA", reduction = 'tsne', order = T)+border()+
  scale_colour_viridis(guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")),

FeaturePlot(Hep59.1.2.seu, features = "nCount_RNA", reduction = 'tsne', order = T, max.cutoff = "q99")+border()+
  scale_colour_viridis(guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")),

FeaturePlot(Hep59.1.2.seu, features = "PVP01-0304700", reduction = "tsne")+border()+ggtitle("PVP01_0304700::liver specific protein 2")+
  border()+scale_colour_viridis(guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")),
align = "hv", nrow = 1

)

# TSNE - nFeatureRNA, nCountRNA, LISP2
ggarrange(
  
Hep59.1.2.seu@meta.data %>% 
    ggplot(aes(x = replicate, fill = day))+
    geom_bar(stat = "count", color = "black")+
    theme_classic()+border()+
    labs(y = "Number of parasites", x = "")+
    scale_x_discrete(labels = c('1' = "Rep 1", '2' = "Rep 2"))+
    scale_fill_manual(values = c("#3C5488FF", "#DC0000FF"))+theme(axis.text = element_text(size=10,color="black"),
                                                                  strip.text = element_text(colour = 'black')),
Hep59.1.2.seu@meta.data %>% 
  as_tibble() %>% 
  ggplot(aes(x = seurat_clusters, fill = seurat_clusters))+
  geom_bar(stat = "count", color = "black")+
  labs(x = "Cluster", y = "Number of parasites")+
  theme(legend.position = "none")+scale_fill_manual(values = c("#00A087FF", "#4DBBD5FF"))+
  facet_grid(replicate~day)+theme_classic2()+border()+theme(axis.text = element_text(size=10,color="black"),
                                                      strip.text = element_text(colour = 'black'))+NoLegend(),

nrow = 1, widths =c (2,3))
```

Save RDS file.
```{r save RDS rep 1 and 2 day 5 and 9 combined}
saveRDS(Hep59.1.2.seu, "outputs/Hep59.1.2.seu.rds")
```

We will now process the hypnozoite-only data.
```{r hypnozoites only}
Hypnozoites<-subset(x = Hep59.1.2.seu, idents = "1")
Hypnozoites$form<-"Hypnozoites"

DefaultAssay(Hypnozoites) <- "SCT"

Hypnozoites <- RunPCA(Hypnozoites , verbose = FALSE)
ElbowPlot(Hypnozoites)
Hypnozoites<- RunUMAP(Hypnozoites, dims = 1:20, verbose = FALSE, min.dist = 0.45 , n.components = 3, umap.method = "umap-learn", seed.use = 7271717)
Hypnozoites<- RunTSNE(Hypnozoites , dims = 1:20, verbose = FALSE, perplexity = 20)

Hypnozoites <- FindNeighbors(Hypnozoites , dims = 1:20, verbose = FALSE)
Hypnozoites <- FindClusters(Hypnozoites, verbose = FALSE, resolution = 0.3, algorithm = 4)
```

Let's visualize the hypnozoite data.
```{r visualize hypnozoite data}
DimPlot(Hypnozoites, group.by =  "day", reduction = "pca")+border()
DimPlot(Hypnozoites, reduction = "umap", group.by = 'day', cols = c('Five' = "#3C5488FF", 'Nine' = "#DC0000FF"), dims = c(1,2))+border()+ggtitle("")
DimPlot(Hypnozoites, label = TRUE, split.by = "replicate", group.by = "day") + NoLegend()+border()
DimPlot(Hypnozoites, label = TRUE, reduction = "umap", dims = c(1,2), split.by = "day") + NoLegend()+border()
DimPlot(Hypnozoites, label = TRUE, split.by = "day", reduction = "umap") + NoLegend()+border()+scale_color_npg()
DimPlot(Hypnozoites, label = TRUE, reduction = "umap")+border()+scale_color_npg()
```

We can now subset the data based on their cluster status.
```{r subset hypnozoites based on cluster status}
Hypnozoites.dormant<-subset(x = Hypnozoites, idents = "1")
Hypnozoites.transition<-subset(x = Hypnozoites, idents = "2")
Hypnozoites.reactivating<-subset(x = Hypnozoites, idents = "3")
```

Save Hypnozoite RDS files
```{r save hypnozoite RDS files}
saveRDS(Hypnozoites, "outputs/Hypnozoites.rds")
saveRDS(Hypnozoites.dormant, "outputs/Hypnozoites.dormant.rds")
saveRDS(Hypnozoites.transition, "outputs/Hypnozoites.transition.rds")
saveRDS(Hypnozoites.reactivating, "outputs/Hypnozoites.activating.rds")
```